\chapter{Differential Equations}

\section{Ordinary Differential Equations}
\subsection{Definitions and Examples}
\begin{definition}[Ordinary Differential Equation]
    An ordinary differential equation (ODE) is an equation that relates a function \( y(t) \) of a single variable \( t \) to its derivatives. The general form of an ODE can be expressed as:
    \[
    F\left(t, y(t), y'(t), y''(t), \ldots, y^{(n)}(t)\right) = 0
    \]
    where \( F \) is a given function and \( n \in \mathbb{N}_+ \) is the order of the ODE.
\end{definition}

\begin{definition}[Solution of an ODE]
    A solution of the differential equation is a function $y: I \to \mathbb{R}$ of class $C^n$ defined on a open interval $I \subset \mathbb{R}$ such that when substituted into the ODE, it satisfies the equation for all $t \in I$.
\end{definition}

\begin{definition}[General Solution]
    The general solution of an ODE is a family of solutions that contains all possible solutions of the ODE. It typically includes arbitrary constants that can be adjusted to fit specific initial conditions or boundary conditions.
\end{definition}

\begin{eg}
    Let $y = y(x)$, $x \in \mathbb{R}$ such that $y' = 0$ for all $x \in \mathbb{R}$. Then $y$ is a constant function, i.e., $y(x) = C$ for some constant $C \in \mathbb{R}$. This is the general solution of the ODE $y' = 0$. A particular solution can be obtained by specifying an initial condition, for example, if we set $y(0) = 5$, then the particular solution is $y(x) = 5$ for all $x \in \mathbb{R}$.
\end{eg}

\begin{eg}
    Let $y'' = 3$ be the ODE. Integrating once, we get $y' = 3x + C_1$ for some constant $C_1 \in \mathbb{R}$. Integrating again, we obtain $y = \frac{3}{2}x^2 + C_1 x + C_2$ for some constant $C_2 \in \mathbb{R}$. Thus, the general solution of the ODE $y'' = 3$ is given by:
    \[
        y(x) = \frac{3}{2}x^2 + C_1 x + C_2
    \]
    where $C_1$ and $C_2$ are arbitrary constants. A particular solution can be found by specifying initial conditions, for example, if we set $y(0) = 2$ and $y'(0) = 1$, we can determine the constants $C_1$ and $C_2$ to find the specific solution that satisfies these conditions.
\end{eg}

\begin{eg}
    Let $y + y' = 0$ be the ODE. We can rearrange this equation to get $y' = -y$. We easily verify that the function $y(x) = Ce^{-x}$, where $C$ is an arbitrary constant, is a solution to this ODE on $\mathbb{R}$. Indeed, we have:
    \[
        y'(x) = -Ce^{-x} = -y(x)
    \]
    which satisfies the ODE. Thus, the general solution of the ODE $y + y' = 0$ is given by:
    \[
        y(x) = Ce^{-x}
    \]
    where $C$ is an arbitrary constant.
\end{eg}

\subsection{Terminology and Notation}
Let $E(x, y, \ldots, y^{(n)}) = 0$ be a differential equation.

\begin{definition}[Order of a Differential Equation]
    The order of a differential equation is the highest order of the derivative that appears in the equation.
\end{definition}
In the given equation, if $y^{(n)}$ is the highest derivative present, then the order of the differential equation is $n$.

\begin{definition}[Linear Differential Equation]
    A differential equation is said to be linear if it can be expressed in the form:
    \[
    a_n(x) y^{(n)} + a_{n-1}(x) y^{(n-1)} + \ldots + a_1(x) y' + a_0(x) y = g(x)
    \]
    where $a_n(x), a_{n-1}(x), \ldots, a_0(x)$ and $g(x)$ are given functions of $x$ that are continuous.
\end{definition}

\begin{definition}[Autonomous Differential Equation]
    A differential equation is called autonomous if it does not explicitly depend on the independent variable.
\end{definition}

\begin{eg}
    Let's classify the following differential equations:
    \begin{citemize}
        \item $(2 \sin x) y + x^2 y' = \cos x$ is a first-order linear non-autonomous ODE.
        \item $y^2 - y'' + y = 0$ is a second-order nonlinear autonomous ODE.
        \item $y''' + 3x^2y = e^x$ is a third-order linear non-autonomous ODE.
        \item $2xy = \sin y + y'$ is a first-order nonlinear non-autonomous ODE.
    \end{citemize}
\end{eg}

\section{Separable Differential Equations}
\begin{definition}[Separable Differential Equation]
    A first-order ordinary differential equation is called separable if it can be expressed in the form:
    \[
        f(y) y' = g(x)
    \]
    where $f: I \to \mathbb{R}$ and $g: J \to \mathbb{R}$ are continuous functions defined on intervals $I$ and $J$ respectively.
\end{definition}
Note that a solution $y$ of a separable differential equation is a function $y: J' \subset J \to I$ of class $C^1$.

\begin{eg}
    Let's come back to the previous equations $y' = -y$. This is a separable ODE since we can rewrite it as:
    \[
        \underbrace{-\frac{1}{y}}_{f(y)}y' = \underbrace{1}_{g(x)} \Implies \frac{1}{y} \frac{dy}{dx} = -1 
    \]
    We then can integrate both side to find the general solution:
    \[
        \int \frac{1}{y} dy = \int -1 dx \Implies \ln |y| + C_1 = -x + C_2 \Implies |y| = e^{-x + C} = \underbrace{e^C}_{= C_3 > 0} e^{-x}
    \]
    where $C = C_2 - C_1$ is an arbitrary constant. Thus:
    \[
        y = \underbrace{\pm C_3}_{\neq 0} e^{-x}
    \]
    but we found in the previous example that $y = 0$ is also a solution of the ODE. Therefore, the general solution of the ODE $y' = -y$ is given by:
    \[
        y(x) = Ce^{-x}
    \]
    where $C$ is an arbitrary constant that can take any real value, including zero.
\end{eg}

\section{Cauchy's Problem}
\begin{definition}[Cauchy's Problem]
    Cauchy's problem, also known as the initial value problem, is a type of problem in the theory of ordinary differential equations. It consists of finding a solution to a given ordinary differential equation that satisfies specified initial conditions. Formally, given an ODE of the form:
    \[
        F\left(t, y(t), y'(t), y''(t), \ldots, y^{(n)}(t)\right) = 0
    \]
    and initial conditions:
    \[
        y(t_0) = y_0, \quad y'(t_0) = y_1, \quad \ldots, \quad y^{(n-1)}(t_0) = y_{n-1}
    \]
    where \( t_0 \) is the initial time and \( y_0, y_1, \ldots, y_{n-1} \) are given constants, the goal is to find a function \( y(t) \) and it's interval of definition that satisfies both the ODE and the initial conditions.
\end{definition}

\begin{definition}[Maximal Solution]
    A solution of an ODE is called maximal if there is no other bigger interval on which the function can be defined as a solution of the ODE.
\end{definition}

\begin{eg}
    Let's solve $y'' = 0$ with $y(0) = 1$ and $y'(0) = 2$. Integrating once, we get $y' = C_1$ for some constant $C_1 \in \mathbb{R}$. Integrating again, we obtain $y = C_1 x + C_2$ for some constant $C_2 \in \mathbb{R}$. Applying the initial conditions, we have:
    \[
        y(0) = C_2 = 1
    \]
    and
    \[
        y'(0) = C_1 = 2
    \]
    Thus, the particular solution that satisfies the initial conditions is given by:
    \[
        y(x) = 2x + 1
    \]
    for all $x \in \mathbb{R}$.
\end{eg}

\begin{eg}
    Let's solve $y'' = 0$ with $y(0) = 3$ and $y(4) = 1$. Integrating once, we get $y' = C_1$ for some constant $C_1 \in \mathbb{R}$. Integrating again, we obtain $y = C_1 x + C_2$ for some constant $C_2 \in \mathbb{R}$. Applying the initial conditions, we have:
    \[
        y(0) = C_2 = 3
    \]
    and
    \[
        y(4) = 4C_1 + C_2 = 1 \Implies C_1 = -\frac{1}{2}
    \]
    Thus, the particular solution that satisfies the initial conditions is given by:
    \[
        y(x) = -\frac{1}{2}x + 3
    \]
    for all $x \in \mathbb{R}$.
\end{eg}

% LEARN: THEOREM SELECTED FOR EXAMS
\begin{theorem}[Existence and Uniqueness Theorem for Separable ODEs]
    Let $f: I \to \mathbb{R}$ be a continuous function such that $f(y) \neq 0$ for all $y \in I$ and $g: J \to \mathbb{R}$ continuous. The for any couple $(x_0 \in J, b_0 \in I)$ the equation:
    \[
        f(y) y' = g(x)
    \]
    has a solution $y: J' \subset J \to I$ verifying the initial condition $y(x_0) = b_0$. Moreover, if $y_1$ and $y_2$ are two solutions verifying the same initial condition, then $y_1 = y_2$ on the intersection of their domains.
\end{theorem}
\begin{proof}
    Let's prove the existence of a solution. Let:
    \[
        F(y) = \int_{b_0}^{y} f(t) dt
    \]
    Then $F$ is differentiable and monotone since $F'(y) = f(y) \neq 0$ on $I$. Thus, $F(I)$ is a open interval and since:
    \[
        b_0 \in I \Implies F(b_0) = 0 \in F(I)
    \]
    Thus $F$ is a bijection of $I$ in a neighbourhood of $0$ and thus it admits an inverse function $F^{-1}$ defined on a neighbourhood of $0$. Let:
    \[
        G(x) = \int_{x_0}^{x} g(t) dt
    \]
    Again, $G$ is differentiable on $J$, $G'(x) = g(x)$ and $G(x_0) = 0$. Let $y(x) = F^{-1}(G(x))$ (defined on a neighbourhood of $x_0$). We have:
    \[
        F(y(x)) = G(x) \Implies F'(y(x)) y'(x) = G'(x) \Implies f(y(x)) y'(x) = g(x)
    \]
    by the definition of $F$ and $G$. Moreover, we have:
    \[
        y(x_0) = F^{-1}(G(x_0)) = F^{-1}(0) = b_0
    \]
    by the definition of $F$ and $G$. Thus, we have found a solution $y$ of the ODE that satisfies the initial condition.
\end{proof}

\begin{eg}
    Let's solve $x^2y' = e^y$ with $y(-2) = 0$. We can rewrite the equation as:
    \[
        \underbrace{e^{-y}}_{f(y)} y' = \underbrace{\frac{1}{x^2}}_{g(x)} \Implies e^{-y} \frac{dy}{dx} = \frac{1}{x^2}
    \]
    we have that $f(y)$ is not equal to $0$ for all $y \in \mathbb{R}$ and $g(x)$ is continuous on $\mathbb{R} \setminus \{0\}$. We can then integrate both sides to find the solution:
    \[
        \int e^{-y} dy = \int \frac{1}{x^2} dx \Implies -e^{-y} = -\frac{1}{x} + C \Implies -y = \ln\left(\frac{1}{x} - C\right)
    \]
    Thus:
    \[
        y(x) = -\ln\left(\frac{1}{x} - C\right)
    \]
    For this solution to exist, we need $\frac{1}{x} - C > 0$, which implies $C < \frac{1}{x}$. Let's find the intervals on which the solution is defined. We have three cases:
    \begin{citemize}
        \item If $C > 0$, then:
        \[
            \frac{1}{x} > C \Implies x > 0 \Implies x < \frac{1}{C} \quad \text{and} \quad x > 0 \Implies x \in \left(0, \frac{1}{C}\right)
        \]
        \item If $C = 0$, then:
        \[
            \frac{1}{x} > 0 \Implies x > 0 \Implies x \in (0, \infty)
        \]
        \item If $C < 0$, then:
        \[
            \frac{1}{x} \cdot \frac{1}{C} < 1 \Implies \begin{cases*}
                x > 0 \Implies x > \frac{1}{C} \Implies x \in (0, \infty) \\
                x < 0 \Implies x < \frac{1}{C} \Implies x \in \left(-\infty, \frac{1}{C}\right)
            \end{cases*}
        \]
    \end{citemize}
    Thus the general solution of the ODE $x^2y' = e^y$ is given by:
    \[
        \begin{cases}
            y = -\ln(\frac{1}{x}) = \ln x, & x \in (0, \infty) \\
            y = -\ln(\frac{1}{x} - C), & x \in \left(0, \frac{1}{C}\right), C > 0 \\
            y = -\ln(\frac{1}{x} - C), & x \in (0, \infty), C < 0 \\
            y = -\ln(\frac{1}{x} - C), & x \in \left(-\infty, \frac{1}{C}\right), C < 0
        \end{cases}
    \]
    Applying the initial condition $y(-2) = 0$, we have:
    \[
        0 = -\ln\left(\frac{1}{-2} - C\right) \Implies \frac{1}{-2} - C = 1 \Implies C = -\frac{3}{2}
    \]
    Thus, the particular solution that satisfies the initial condition is given by:
    \[
        y(x) = -\ln\left(\frac{1}{x} + \frac{3}{2}\right)
    \]
    which is defined on the interval $\left(-\infty, -\frac{2}{3}\right)$.
\end{eg}

\begin{eg}
    Let's solve:
    \[
        \frac{y'(x)}{y^2(x)} = 1
    \]
    with $y(0) = b_0 \in \mathbb{R}^*$. We can rewrite the equation as:
    \[
        \underbrace{\frac{1}{y^2}}_{f(y)} y' = \underbrace{1}_{g(x)} \Implies \frac{1}{y^2} \frac{dy}{dx} = 1
    \]
    we have that $f(y)$ is not equal to $0$ for all $y \in \mathbb{R}^*$ and $g(x)$ is continuous on $\mathbb{R}$. We can then integrate both sides to find the solution:
    \[
        \int \frac{1}{y^2} dy = \int 1 dx \Implies -\frac{1}{y} = x + C \Implies y = -\frac{1}{x + C}
    \]
    Applying the initial condition $y(0) = b_0$, we have:
    \[
        b_0 = -\frac{1}{C} \Implies C = -\frac{1}{b_0}
    \]
    Which is the general solution defined on the intervals $\left(- \infty, -C\right)$ and $\left(-C, \infty\right)$. The particular solution that satisfies the initial condition is given by:
    \[
        y(x) = -\frac{1}{x - \frac{1}{b_0}} = \frac{1}{\frac{1}{b_0} - x}
    \]
    To find the maximal interval on which this solution is defined, we need to find the values of $x$ such that $\frac{1}{b_0} - x \neq 0$, which implies $x \neq \frac{1}{b_0}$. Thus, the maximal interval on which the solution is defined is given by:
    \[
        \begin{cases}
            \left(-\infty, \frac{1}{b_0}\right), & b_0 > 0 \\
            \left(\frac{1}{b_0}, \infty\right), & b_0 < 0
        \end{cases}
    \]
    Note that we chose the intervals to include the point $x = 0$ in the maximal interval since it is the point at which the initial condition is defined.
\end{eg}

\section{First Order Linear Differential Equations}
\begin{definition}[First Order Linear Differential Equation]
    Let $I \subset \mathbb{R}$ be an open interval. A first order linear differential equation is an equation of the form:
    \[
        y' + p(x) y = f(x)
    \]
    where $p, f: I \to \mathbb{R}$ are continuous functions defined on $I$. A solution of this equation is a function $y: I \to \mathbb{R}$ of class $C^1$ that satisfies the equation.
\end{definition}

\begin{definition}[Associated Homogeneous Equation]
    The associated homogeneous equation of the first order linear differential equation $y' + p(x) y = f(x)$ is the equation:
    \[
        y' + p(x) y = 0
    \]
\end{definition}

\subsection{Principal of Superposition}
\begin{theorem}
    Let $p,f_1, f_2: I \to \mathbb{R}$ be continuous functions defined on an open interval $I$. If $v_1$ is a solution of:
    \[
        y' + p(x)y = f_1(x)
    \]
    and $v_2$ is a solution of:
    \[
        y' + p(x)y = f_2(x)
    \]
    Then for all $C_1, C_2 \in \mathbb{R}$, the function $v = C_1 v_1 + C_2 v_2$ is a solution of:
    \[
        y' + p(x)y = C_1 f_1(x) + C_2 f_2(x)
    \]
\end{theorem}
\begin{proof}
    We have:
    \begin{align*}
        v' + p(x)v & = C_1 v_1' + C_2 v_2' + p(x)(C_1 v_1 + C_2 v_2) \\
        & = C_1 (v_1' + p(x) v_1) + C_2 (v_2' + p(x) v_2) \\
        & = C_1 f_1(x) + C_2 f_2(x)
    \end{align*}
    where we used the fact that $v_1$ and $v_2$ are solutions of their respective equations.
\end{proof}

\subsection{Variation of Parameters}
We want to find a particular solution of the equation:
\[
    y' + p(x) y = f(x)
\]
of the form $v(x) = C(x) e^{-P(x)}$, where $P(x)$ is a primitive of $p(x)$ and $C(x)$ is a function to be determined. We have:
\begin{align*}
    v' + p(x) v & = C'(x) e^{-P(x)} + C(x) \left(-P'(x) e^{-P(x)} + p(x) e^{-P(x)}\right) \\
    & = C'(x) e^{-P(x)} \\
    & = f(x)
\end{align*}
which implies that $C'(x) = f(x) e^{P(x)}$. Integrating both sides, we get:
\[
    C(x) = \int f(x) e^{P(x)} dx
\]Thus, a particular solution of the equation $y' + p(x) y = f(x)$ is given by:
\[
    v(x) = e^{-P(x)} \int f(x) e^{P(x)} dx
\]
where $P$ is a primitive of $p$.

\begin{theorem}
    Let $p,f: I \to \mathbb{R}$ be continuous functions. If $v_0$ is a particular solution of $y' + p(x)y = f(x)$, then the general solution of this equation is given by:
    \[
        v(x) = v_0(x) + Ce^{-P(x)}
    \]
    where $P$ is a primitive of $p$ on $I$ and $C \in \mathbb{R}$.
\end{theorem}
\begin{proof}
    Let $v_1$ be an arbitrary solution of the equation 
    \[
        y' + p(x)y = f(x)
    \] 
    Let's show that there exists $C \in \mathbb{R}$ such that:
    \[
        v_1(x) = v_0(x) + Ce^{-P(x)}
    \]
    where $v_0$ is a particular solution of the equation. We have by the principal of superposition that $v_1 - v_0$ is a solution of the associated homogeneous equation $y' + p(x) y = 0$. Thus, there exists $C \in \mathbb{R}$ such that: 
    \[
        v_1 - v_0 = Ce^{-P(x)} \Implies v_1(x) = v_0(x) + Ce^{-P(x)}
    \]
    Since $v_1$ is an arbitrary solution of the equation, we have shown that any solution of the equation can be written in the form $v(x) = v_0(x) + Ce^{-P(x)}$ for some $C \in \mathbb{R}$ and thus by definition, that the general solution of the equation is given by $v(x)$.
\end{proof}

\subsection{Solving First Order Linear Differential Equations}
To solve a first order linear differential equation of the form $y' + p(x) y = f(x)$, we can follow these steps:
\begin{citemize}
    \item Find the general solution of the associated homogeneous equation $y' + p(x) y = 0$.
    \item Find a particular solution of the original equation $y' + p(x) y = f(x)$ using the method of variation of parameters.
    \item Combine the general solution of the homogeneous equation and the particular solution to get the general solution of the original equation.
\end{citemize}

\begin{eg}
    Let's find the general solution of the equation:
    \[
        y' + \frac{1}{x} y = \frac{1}{x + 1}
    \]
    on the interval $(0, \infty)$. First, we find the general solution of the associated homogeneous equation $y' + \frac{1}{x} y = 0$. We can rewrite this equation as:
    \[
        \frac{1}{y} y' = -\frac{1}{x}
    \]
    Integrating both sides, we get:
    \[
        \int \frac{1}{y} dy = \int -\frac{1}{x} dx \Implies \ln |y| = -\ln |x| + K \Implies |y| = e^{- \ln |x| + K} = \underbrace{e^K}_{= C > 0} \frac{1}{x}
    \]
    Thus, the general solution of the homogeneous equation is given by:
    \[
        y(x) = \frac{C}{x}
    \]
    where $C$ is an arbitrary constant. Next, we find a particular solution of the original equation using the method of variation of parameters. We have:
    \[
        C(x) = \int f(x) e^{P(x)} dx = \int \frac{1}{x + 1} e^{\ln x} dx = \int \frac{x}{x + 1} dx
    \]
    We can rewrite the integrand as:
    \[
        \frac{x}{x + 1} = \frac{(x + 1) - 1}{x + 1} = 1 - \frac{1}{x + 1}
    \]
    Thus, we have:
    \[
        C(x) = \int \left(1 - \frac{1}{x + 1}\right) dx = \int 1 dx - \int \frac{1}{x + 1} dx = x - \ln |x + 1| = x - \ln (x + 1)
    \]
    Thus, a particular solution of the original equation is given by:
    \[
        y(x) = C(x) e^{-P(x)} = (x - \ln (x + 1)) e^{-\ln x} = 1 - \frac{\ln (x + 1)}{x}
    \]
    We can verify that this is indeed a solution of the original equation by substituting it back into the equation:
    \begin{align*}
        y' + \frac{1}{x} y & = \left(1 - \frac{\ln (x + 1)}{x}\right)' + \frac{1}{x} \left(1 - \frac{\ln (x + 1)}{x}\right) \\
        & = \ldots \\
        & = \frac{1}{x + 1}
    \end{align*}
    Thus, the general solution of the original equation is given by:
    \[
        y(x) = 1 - \frac{\ln (x + 1)}{x} + \frac{C}{x}
    \]
    where $C$ is an arbitrary constant.
\end{eg}

\begin{eg}
    Let's find the general solution of the equation:
    \[
        y' - \frac{2}{x}y = x^3 + 1
    \]
    First, let's find the interval of definition of the functions $p(x) = -\frac{2}{x}$ and $f(x) = x^3 + 1$. Since $p(x) = -\frac{2}{x}$ is undefined at $x = 0$, the interval of definition is either $(-\infty, 0)$ or $(0, \infty)$. Let's compute $P(x)$, a primitive of $p(x)$:
    \[
        P(x) = \int -\frac{2}{x} dx = -2 \ln |x| = -\ln \left(x^2\right)
    \]
    Thus, the general solution of the associated homogeneous equation $y' - \frac{2}{x} y = 0$ is given by:
    \[
        y(x) = Ce^{- (-\ln \left(x^2\right))} = Cx^2
    \]
    Let's check that this is indeed a solution of the homogeneous equation:
    \[
        y' - \frac{2}{x} y = 2Cx - \frac{2}{x} Cx^2 = 0
    \]
    Next, we find a particular solution of the original equation using the method of variation of parameters. We have:
    \[
        \int f(x) e^{P(x)} dx = \int (x^3 + 1) e^{-\ln \left(x^2\right)} dx = \int (x^3 + 1) \frac{1}{x^2} dx = \int \left(x + \frac{1}{x^2}\right) dx = \frac{x^2}{2} - \frac{1}{x}
    \]
    Thus, a particular solution of the original equation is given by:
    \[
        y(x) = \left(\frac{x^2}{2} - \frac{1}{x}\right) e^{\ln \left(x^2\right)} = \frac{1}{2} x^4 - x
    \]
    We can verify that this is indeed a solution of the original equation by substituting it back into the equation:
    \begin{align*}
        y' - \frac{2}{x} y & = \left(\frac{1}{2} x^4 - x\right)' - \frac{2}{x} \left(\frac{1}{2} x^4 - x\right) \\
        & = 2x^3 - 1 - \frac{2}{x} \left(\frac{1}{2} x^4 - x\right) \\
        & = 2x^3 - 1 - \left(x^3 - 2\right) \\
        & = x^3 + 1
    \end{align*}
    Thus, the general solution of the original equation is given by:
    \[
        y(x) = \frac{1}{2} x^4 - x + Cx^2
    \]
    where $C$ is an arbitrary constant, for all $x \in (-\infty, 0)$ or for all $x \in (0, \infty)$.
\end{eg}
Note that if, instead, we were considering the equation $xy' - 2y = x^4 + x$, the general solution would be given by:
\[
    y(x) = \begin{cases}
        C_1 x^2 + \frac{1}{2}x^4 - x, & x \geq 0 \\
        C_2 x^2 + \frac{1}{2}x^4 - x, & x < 0 \\
    \end{cases}
\]
for all $x \in \mathbb{R}$, where $C_1$ and $C_2$ are arbitrary constants. This is because the function is defined at $x = 0$ and it's derivative at $x = 0$ is $-1$ for all $C \in \mathbb{R}$. Thus, we can choose any value of $C$ for the solution defined on the interval $(-\infty, 0)$ and any value of $C$ for the solution defined on the interval $(0, \infty)$.

\begin{eg}
    Let's solve the equation:
    \[
        y' - (\tan x)y = \cos x
    \]
    with the initial condition $y(0) = 3$. First, let's find the interval of definition of the functions $p(x) = -\tan x$ and $f(x) = \cos x$. Since $\tan x$ is undefined at $x = (2k + 1) \frac{\pi}{2}$ for all $k \in \mathbb{Z}$, we take the interval of definition that contains $0$, which is $I = \left(-\frac{\pi}{2}, \frac{\pi}{2}\right)$. Let's compute $P(x)$, a primitive of $p(x)$:
    \[
        P(x) = \int -\tan x dx = \ln |\cos x| = \ln (\cos x)
    \]
    Thus, the general solution of the associated homogeneous equation $y' - (\tan x) y = 0$ is given by:
    \[
        y(x) = Ce^{-\ln (\cos x)} = \frac{C}{\cos x}
    \]
    Let's check that this is indeed a solution of the homogeneous equation:
    \[
        y' - (\tan x) y = \frac{C \sin x}{\cos^2 x} - \frac{C \sin x}{\cos^2 x} = 0
    \]
    Next, we find a particular solution of the original equation using the method of variation of parameters. We have:
    \[
        \int f(x) e^{P(x)} dx = \int \cos x e^{\ln (\cos x)} dx = \int \cos^2 x dx
    \]
    We can rewrite the integrand as:
    \[
        \cos^2 x = \frac{1 + \cos (2x)}{2}
    \]
    Thus, we have:
    \[
        \int \cos^2 x dx = \int \frac{1 + \cos (2x)}{2} dx = \frac{x}{2} + \frac{\sin (2x)}{4}
    \]
    Thus, a particular solution of the original equation is given by:
    \[
        y(x) = \left(\frac{x}{2} + \frac{\sin (2x)}{4}\right) e^{-\ln (\cos x)} = \left(\frac{x}{2} + \frac{\sin (2x)}{4}\right) \frac{1}{\cos x} = \frac{x}{2 \cos x} + \frac{\sin x}{2}
    \]
    We can verify that this is indeed a solution of the original equation by substituting it back into the equation:
    \begin{align*}
        y' - (\tan x) y & = \left(\frac{x}{2 \cos x} + \frac{\sin x}{2}\right)' - (\tan x) \left(\frac{x}{2 \cos x} + \frac{\sin x}{2}\right) \\
        & = \ldots \\
        & = \cos x
    \end{align*}
    Thus, the general solution of the original equation is given by:
    \[
        y(x) = \frac{x}{2 \cos x} + \frac{\sin x}{2} + \frac{C}{\cos x}
    \]
    where $C$ is an arbitrary constant, for all $x \in \left(-\frac{\pi}{2}, \frac{\pi}{2}\right)$. Applying the initial condition $y(0) = 3$, we have:
    \[
        3 = \frac{0}{2 \cdot 1} + \frac{0}{2} + \frac{C}{1} \Implies C = 3
    \]
    Thus, the particular solution that satisfies the initial condition is given by:
    \[
        y(x) = \frac{x}{2 \cos x} + \frac{\sin x}{2} + \frac{3}{\cos x}
    \]
    for all $x \in \left(-\frac{\pi}{2}, \frac{\pi}{2}\right)$.
\end{eg}

\section{Linear Differential Equations of Second Order}
\begin{definition}[Linear Differential Equation of Second Order]
    Let $I \subset \mathbb{R}$ be an open interval. A linear differential equation of second order is an equation of the form:
    \[
        y'' + p(x) y' + q(x) y = f(x)
    \]
    where $p, q, f: I \to \mathbb{R}$ are continuous functions defined on $I$.
\end{definition}
Note that again, if $f(x) = 0$ for all $x \in I$, (i.e. the equation is $y'' + p(x) y' + q(x) y = 0$) then the equation is called homogeneous, otherwise it is called non-homogeneous.

\begin{eg}
    Let's solve the equation:
    \[
        y'' = 5
    \]
    Integrating once, we get $y' = 5x + C_1$ for some constant $C_1 \in \mathbb{R}$. Integrating again, we obtain $y = \frac{5}{2} x^2 + C_1 x + C_2$ for some constant $C_2 \in \mathbb{R}$. Thus, the general solution of the equation is given by:
    \[
        y(x) = \frac{5}{2} x^2 + C_1 x + C_2
    \]
    for all $x \in \mathbb{R}$, where $C_1$ and $C_2$ are arbitrary constants.
\end{eg}

\subsection{Equation with Constant Coefficients}
\begin{definition}[Linear Differential Equation of Second Order with Constant Coefficients]
    A linear differential equation of second order with constant coefficients is an equation of the form:
    \[
        y'' + py' + qy = 0
    \]
    where $p, q \in \mathbb{R}$ are constants.
\end{definition}

\begin{definition}[Characteristic Polynomial]
    The characteristic polynomial of the linear differential equation of second order with constant coefficients $y'' + py' + qy = 0$ is the polynomial:
    \[
        P(\lambda) = \lambda^2 + p\lambda + q = 0
    \]
\end{definition}
Thus to solve a linear differential equation of second order with constant coefficients, we need to find the roots of the characteristic polynomial:
\[
    (\lambda - a) (\lambda - b) = \lambda^2 - (a + b) \lambda + ab = 0
\]
where $a,b \in \mathbb{C}$. Thus the equation can be rewritten as:
\[
    y''(x) - (a + b)y'(x) + aby(x) = 0 \Implies \underbrace{(y'(x) - ay(x))'}_{z'(x)} - b\underbrace{(y'(x)-ay(x))}_{z(x)} = 0
\]
Thus, we have $z'(x) - bz(x) = 0$, which is a first order linear differential equation whose general solution is given by $z(x) = C e^{bx}$ for some constant $C \in \mathbb{R}$. Thus, we have:
\[
    y'(x) - ay(x) = C e^{bx} \Implies y'(x) - ay(x) = C e^{bx}
\]
which is a first order linear differential equation. We can solve it using the method of variation of parameters. We have:
\[
    P(x) = \int -a dx = -ax
\]
Thus, the general solution of the associated homogeneous equation $y' - ay = 0$ is given by:
\[
    y(x) = C_1e^{-P(x)} = C_1e^{ax}
\]
Next, we find a particular solution of the original equation using the method of variation of parameters. We have:
\[
    c(x) = \int f(x) e^{P(x)} dx = \int C e^{bx} e^{-ax} dx = \int C e^{(b - a)x} dx = \begin{cases}
    \frac{C}{b - a} e^{(b - a)x}, & a \neq b \\
    Cx, & a = b
    \end{cases}
\]
Thus, a particular solution of the original equation is given by:
\[
    y(x) = c(x) e^{-P(x)} = \begin{cases}
    \frac{C}{b - a} e^{bx} = K e^{bx}, & a \neq b \\
    Cx e^{ax}, & a = b
    \end{cases}
\]
where $K = \frac{C}{b - a}$ is an arbitrary constant. Thus, the general solution of the original equation is given by:
\[
    y(x) = \begin{cases}
    C_1 e^{ax} + K e^{bx}, & a \neq b \\
    C_1 e^{ax} + C x e^{ax}, & a = b
    \end{cases}
\]
where $C$ and $C_1$ are arbitrary constants. Note that if $a,b \notin \mathbb{R}$, then $b = \overline{a}$ since $p,q \in \mathbb{R}$. Let $a = \alpha + i\beta$ where $\beta \neq 0$, then we have:
\[
    y(x) = Ce^{ax} + \overline{C} e^{\overline{a}x}
\]
where $C = \frac{1}{2}(C_2 - i C_3)$, we get:
\[
    y(x) = \frac{1}{2} (C_2 - i C_3) e^{\alpha x} e^{i \beta x} + \frac{1}{2} (C_2 + i C_3) e^{\alpha x} e^{-i \beta x} = e^{\alpha x} \left(C_2 \cos (\beta x) + C_3 \sin (\beta x)\right)
\]
where $C_2$ and $C_3$ are arbitrary constants. Thus, the general solution of the original equation is given by:
\[
    y(x) = \begin{cases}
    C_1 e^{ax} + K e^{bx}, & a \neq b, a,b \in \mathbb{R} \\
    C_1 e^{ax} + C x e^{ax}, & a = b, a,b \in \mathbb{R} \\
    e^{\alpha x} \left(C_2 \cos (\beta x) + C_3 \sin (\beta x)\right), & a,b = \alpha \pm i \beta \notin \mathbb{R}
    \end{cases}
\]
where $C, C_1, C_2$ and $C_3$ are arbitrary constants.

\begin{eg}
    Let's find the general solution of the equation:
    \[
        y'' + 9y = 0
    \]
    The characteristic polynomial of this equation is given by:
    \[
        P(\lambda) = \lambda^2 + 9 = 0
    \]
    which has the roots $3i$ and $-3i$. Thus, the general solution of the equation is given by:
    \[
        y(x) = e^{0 \cdot x} \left(C_1 \cos (3x) + C_2 \sin (3x)\right) = C_1 \cos (3x) + C_2 \sin (3x)
    \]
    where $C_1$ and $C_2$ are arbitrary constants, for all $x \in \mathbb{R}$. To verify that this is indeed a solution of the equation, we can substitute it back into the equation:
    \begin{align*}
        y'' + 9y & = -9C_1 \cos (3x) - 9C_2 \sin (3x) + 9C_1 \cos (3x) + 9C_2 \sin (3x) \\
        & = 0
    \end{align*}
\end{eg}

\begin{eg}
    Let's solve the equation:
    \[
        y'' - 6y' + 9y = 0
    \]
    The characteristic polynomial of this equation is given by:
    \[
        P(\lambda) = \lambda^2 - 6\lambda + 9 = 0
    \]
    which has the root $3$ with multiplicity $2$. Thus, the general solution of the equation is given by:
    \[
        y(x) = C_1 e^{3x} + C_2 x e^{3x}
    \]
    where $C_1$ and $C_2$ are arbitrary constants, for all $x \in \mathbb{R}$.
\end{eg}

\subsection{General Homogeneous Linear Differential Equations of Second Order}
Let's consider the general homogeneous linear differential equation of second order:
\[
    y'' + p(x) y' + q(x) y = 0
\]
where $p, q: I \to \mathbb{R}$ are continuous functions defined on an open interval $I$. \\
Remark that the principal of superposition still holds for the general homogeneous linear differential equation of second order. Thus, if $y_1$ and $y_2$ are two solutions of the equation, then for all $C_1, C_2 \in \mathbb{R}$, the function $y = C_1 y_1 + C_2 y_2$ is also a solution of the equation.
\begin{theorem}
    A linear differential equation of the second order has a unique solution $y(x)$ of class $C^2$ that satisfies the initial conditions $y(x_0) = y_0$ and $y'(x_0) = y_1$ for all $x_0 \in I$ and for all $y_0, y_1 \in \mathbb{R}$.
\end{theorem}
By the theorem, we have the the vector space of solutions of the equation is of dimension $2$. There is no known formula to find the general solution of the equation, but if we have one solution $v_1(x) \neq 0$ of the equation, we can find a second solution $v_2$ that is linearly independent of $v_1$ using the following formula:
\[
    v_2(x) = c(x) v_1(x)
\]
such that $c(x)$ is not a constant function. We have:
\begin{align*}
    v_2' & = c'(x) v_1 + c(x) v_1' \\
    v_2'' & = c''(x) v_1 + 2 c'(x) v_1' + c(x) v_1''
\end{align*}
Substituting $v_2$ into the equation, we get:
\[
    c''(x) v_1(x) + 2c'(x) v_1'(x) + p(x) c'(x) v_1(x) = 0
\]
Let's suppose that $v_1(x) \neq 0$ for all $x \in I$ and that $c'(x) \neq 0$ for all $x \in I$, then we have:
\[
    \frac{c''(x)}{c'(x)} = -p(x)-2\frac{v_1'(x)}{v_1(x)}
\]
Integrating both sides, we get:
\[
    \ln |c'(x)| = -\int p(x) dx - 2\int \frac{v_1'(x)}{v_1(x)} dx = \ln (e^{-P(x)}) - 2 \ln |v_1(x)| + \ln K = \ln \left(\frac{Ke^{-P(x)}}{v_1^2(x)}\right)
\]
where $P$ is a primitive of $p$ and $K$ is an arbitrary constant. Thus, we have:
\[
    c'(x) = \pm \frac{Ke^{-P(x)}}{v_1^2(x)} = \frac{K_2e^{-P(x)}}{v_1^2(x)} \Implies c(x) =\int K_2 \frac{e^{-P(x)}}{v_1^2(x)} dx + K_3
\]
where $K_2$ and $K_3$ are arbitrary constants. Thus, a second solution of the equation that is linearly independent of $v_1$ is given by:
\[
    v_2(x) = c(x) v_1(x) = v_1(x) \int\frac{e^{-P(x)}}{v_1^2(x)} dx
\]
where $K_3 = 0$ and $K_2 = 1$.

\begin{eg}
    Let's find a second solution to the previous equation $y'' - 6y' + 9y = 0$ given that $v_1(x) = e^{3x}$ is a solution of the equation. We have:
\[    P(x) = \int -6 dx = -6x
\]
We have:
\[
    c(x) = \int \frac{e^{-P(x)}}{v_1^2(x)} dx = \int \frac{e^{6x}}{e^{6x}} dx = \int 1 dx = x
\]
Thus, a second solution of the equation that is linearly independent of $v_1$ is given by:
\[
    v_2(x) = c(x) v_1(x) = x e^{3x}
\]
and the general solution of the equation is given by:
\[
    y(x) = C_1 e^{3x} + C_2 x e^{3x}
\]
where $C_1$ and $C_2$ are arbitrary constants, for all $x \in \mathbb{R}$.
\end{eg}